---
title: "Quiz 170227"
author: "coop711"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("pander", repos = "https://cran.rstudio.com")
# install.packages("xlsx", repos = "https://cran.rstudio.com")
library(xlsx)
library(pander)
library(knitr)
panderOptions('table.split.table', Inf)
panderOptions('table.alignment.rownames', 'left')
panderOptions('table.alignment.default', 'right')
options(width=180)
```

```{r, data, echo = FALSE, results='hide'}
quiz0227 <- read.xlsx("../data/quiz0227.xlsx", 1, startRow = 2, endRow = 155, colIndex = c(3:7, 9:14), colClasses = c(rep("character", 11)), stringsAsFactors = FALSE)
str(quiz0227)
names(quiz0227) <- c("dept", "id", "name", "year", "e.mail", "cell.no", "Q1", "Q2", "Q3", "Q4", "group")
quiz0227$dept <- factor(quiz0227$dept)
quiz0227$year <- factor(quiz0227$year)
quiz0227$group <- factor(quiz0227$group, levels = c("Red", "Black"))
quiz0227$Q1 <- factor(quiz0227$Q1, levels = c("가", "나", "다", "라"), labels = c("통계학", "통계", "자료", "통계청"))
quiz0227$Q2 <- factor(quiz0227$Q2, levels = c("가", "나", "다", "라"), labels = c("통계학", "통계", "자료", "통계청"))
quiz0227$Q3 <- factor(quiz0227$Q3, levels = c("가", "나", "다", "라"), labels = c("통계학", "통계", "자료", "통계청"))
quiz0227$Q4 <- factor(quiz0227$Q4, levels = c("가", "나", "다", "라", "마"), labels = c("매우 공감", "다소 공감", "다소 비공감", "매우 비공감", "모름/무응답"))
quiz0227$Q4.2 <- quiz0227$Q4
levels(quiz0227$Q4.2) <- c(rep("공감", 2), rep("비공감", 2), "모름/무응답")
str(quiz0227)
```

```{r, merge, echo = FALSE, results = 'hide'}
load("./red_and_black_170227_data.rda")
str(class_roll)
quiz0227.m <- merge(quiz0227, class_roll, by = c("name", "dept", "id", "year", "e.mail", "cell.no"), all = TRUE)
str(quiz0227.m)
```

### Red and Black 분류 집계

#### 출석부와 퀴즈 응답 전체 집계

```{r, regroup, echo = FALSE, results = 'hide'}
group.x.na <- is.na(quiz0227.m$group.x)
# quiz0227.m$group <- ifelse(group.x.na, quiz0227.m$group.y, quiz0227.m$group.x)
quiz0227.m$group <- factor(ifelse(group.x.na, quiz0227.m$group.y, quiz0227.m$group.x), labels = c("Red", "Black"))
kable(quiz0227.m[, c("group.x", "group.y", "group")], col.names = c("퀴즈", "출석부", "병합"))
```

#### 교차표

```{r, cross table, echo = FALSE}
reduced <- quiz0227.m$group.x == quiz0227.m$group.y
quiz0227.m <- quiz0227.m[reduced, ]
tbl.q <- table("출석부" = quiz0227.m$group.y, "퀴즈" = quiz0227.m$group.x, useNA = "ifany")
tbl.q.a <- addmargins(tbl.q)
colnames(tbl.q.a)[3:4] <- c("결석", "계")
rownames(tbl.q.a)[3:4] <- c("추가", "계")
pander(tbl.q.a)
```

### Quiz 응답 비교

#### Dewey as good as elected, statistics convince Roper

##### Roper(Counts) 

```{r, Roper, echo = FALSE}
tbl.q1 <- table(quiz0227.m$group, quiz0227.m$Q1, useNA = "no")
tbl.q1.a <- addmargins(tbl.q1)
colnames(tbl.q1.a)[5] <- "계"
rownames(tbl.q1.a)[3] <- "계"
pander(tbl.q1.a)
```

##### Roper(%)

```{r, Roper in percentage, echo = FALSE}
tbl.q1.a.p <- prop.table(tbl.q1.a[3, -5])
tbl.q1.a.p.sum <- c(tbl.q1.a.p, "계" = sum(tbl.q1.a.p))
pander(round(tbl.q1.a.p.sum*100, digits = 1), caption="Roper")
```

#### Statistics is the science of learning from data, ...

##### ASA(Counts) 

```{r, ASA, echo =FALSE}
tbl.q2 <- table(quiz0227.m$group, quiz0227.m$Q2, useNA = "no")
tbl.q2.a <- addmargins(tbl.q2)
colnames(tbl.q2.a)[5] <- "계"
rownames(tbl.q2.a)[3] <- "계"
pander(tbl.q2.a)
```

##### ASA(%)

```{r, ASA in percentage, echo = FALSE}
tbl.q2.a.p <- prop.table(tbl.q2.a[3, -5])
tbl.q2.a.p.sum <- c(tbl.q2.a.p, "계" = sum(tbl.q2.a.p))
pander(round(tbl.q2.a.p.sum*100, digits = 1), caption = "ASA")
```

#### How to lie with statistics

##### D.Huff(Counts) 

```{r, D.Huff, echo = FALSE}
tbl.q3 <- table(quiz0227.m$group, quiz0227.m$Q3, useNA = "no")
tbl.q3.a <- addmargins(tbl.q3)
colnames(tbl.q3.a)[5] <- "계"
rownames(tbl.q3.a)[3] <- "계"
pander(tbl.q3.a)
```

##### D.Huff(%)

```{r, D.Huff in percentage, echo = FALSE}
tbl.q3.a.p <- prop.table(tbl.q3.a[3, -5])
tbl.q3.a.p.sum <- c(tbl.q3.a.p, "계" = sum(tbl.q3.a.p))
pander(round(tbl.q3.a.p.sum*100, digits = 1), caption = "D.Huff")
```

#### THAAD Question

##### Positive answers are expected to be more frequent in "Red" group, which used lengthy explanation in question.

```{r, THAAD(5 possible answers), echo = FALSE}
tbl.q4 <- table(quiz0227.m$group, quiz0227.m$Q4, useNA = "no")
tbl.q4.a <- addmargins(tbl.q4)
colnames(tbl.q4.a)[6] <- "계"
rownames(tbl.q4.a)[3] <- "계"
pander(tbl.q4.a[-3, ])
```

##### {"매우 공감", "다소 공감" => "공감"}, {"다소 비공감", "매우 비공감" => "비공감"}

```{r, THAAD(3 possible answers), echo = FALSE}
tbl.q4.2 <- table(quiz0227.m$group, quiz0227.m$Q4.2, useNA = "no")
tbl.q4.2.a <- addmargins(tbl.q4.2)
colnames(tbl.q4.2.a)[4] <- "계"
rownames(tbl.q4.2.a)[3] <- "계"
pander(tbl.q4.2.a[-3, ])
```

##### THAAD(%)

```{r, THAAD(%), echo = FALSE}
tbl.q4.p <- prop.table(tbl.q4, margin = 1)
tbl.q4.p.sum <- cbind(tbl.q4.p, rowSums(tbl.q4.p))
colnames(tbl.q4.p.sum)[6] <- "계"
pander(round(tbl.q4.p.sum*100, digits = 1), caption = "THAAD(5문항, %)")
tbl.q4.2.p <- prop.table(tbl.q4.2, margin = 1)
tbl.q4.2.p.sum <- cbind(tbl.q4.2.p, rowSums(tbl.q4.2.p))
colnames(tbl.q4.2.p.sum)[4] <- "계"
pander(round(tbl.q4.2.p.sum*100, digits = 1), caption = "THAAD(3문항, %)")
```

<!--
```{r, save, echo = FALSE}
save.image("../R/quiz170227.RData")
```
-->